---
title: "ingest_data"
author: "<h5>Author: <i>Brian M. Schilder</i></h5>" 
date: "<h5>Most recent update: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h5>"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ingest_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(scKirby)
```

# Intro 

`ingest_data()` runs three main steps:

1. **Read**: Automatically infers the file/object type and loads it (sub-function: `read_data()`).  
2. **Convert**: Converts it to the desired file/object type (sub-function: `to_<format>`). 
3. **Save**: Saves the converted file/object (sub-function: `save_data()`).

# Examples 

## Ingest expression matrix  

### As `SingleCellExperiment`  

```{r expression matrix}
data("example_seurat")

sce <- ingest_data(obj=example_seurat@assays$RNA@counts)
```

### As `Seurat` 

```{r}
seurat <- ingest_data(obj=example_seurat@assays$RNA@counts, 
                      output_type = "Seurat") 
```



## Ingest list 

`scKirby` can ingest a named list (i.e. `list(exp=..., annot=...)`) with the following items:  

- `exp`: Expression matrix with *rows/genes x cols/cells*. 
Can be a variety of matrix classes, including dense or sparse.  

- `annot`: Cell annotation `data.frame` with one cell per row. 
`rownames(annot)` should be the same as `colnames(exp)`. 

This happens to be the format that the example data in [`EWCE`](https://github.com/NathanSkene/EWCE) uses, but any user-supplied data will work. 

### As `SingleCellExperiment`  

```{r EWCElist} 
data("example_EWCElist")

sce <- ingest_data(obj=example_EWCElist)
```

### As `Seurat` 

```{r} 
seurat <- ingest_data(obj=example_EWCElist, 
                      output_type = "Seurat")
```



## Ingest Seurat 

### As `SingleCellExperiment`  

In-memory

```{r Seurat}
data("example_seurat")

sce <- ingest_data(obj=example_seurat)
```

### Ingest H5Seurat  

On-disk 

```{r H5Seurat} 
library(SeuratDisk)
data("example_seurat")

SaveH5Seurat(example_seurat, filename = "./pbmc_small.h5Seurat", overwrite = T)
seurat <- ingest_data(obj="./pbmc_small.h5Seurat")
```


## Ingest HDF5 SingleCellExperiment 

### As `SingleCellExperiment`  

```{r HDF5}
data("example_sce")

sce <- HDF5Array::saveHDF5SummarizedExperiment(example_sce, dir = "./pbmc_small_h5", replace=T)
## Read in the sce object directly
sce <- ingest_data(obj=sce)

## Read it from disk
sce_dir <- dirname(sce_filepath(sce))
sce <- ingest_data(obj=sce_dir)
```

### As `Seurat` 

```{r}
sce <- HDF5Array::saveHDF5SummarizedExperiment(example_sce, dir = "./pbmc_small_h5", replace=T)
## Read in the sce object directly
seurat <- ingest_data(obj=sce,
                      output_type = "Seurat") 
```



## Ingest AnnData 

For info on setting up anndata (e.g. with `conda`), see 

Note that some objects need to be loaded via functions instead of `data(<name>)` 
(e.g. `example_anndata()` and `example_loom()`). 
This is because file types like `loom` and `anndata` must be stored on-disk.  

### As `SingleCellExperiment`  

```{r AnnData}
### Set condaenv= to the name of a conda env you've made 
reticulate::use_condaenv(condaenv = "echoR")

# Convert Seurat object to AnnData for example data
adata <- example_anndata()
## In memory
sce <- ingest_data(obj=adata)

## On disk
adata$write_h5ad(filename = "./pbmc_small.h5ad")
sce <- ingest_data(obj = "./pbmc_small.h5ad")
```

### As `Seurat` 

```{r}
seurat <- ingest_data(obj=adata, 
                      output_type = "Seurat")
```





## Ingest loom

From [loomR](https://github.com/mojaveazure/loomR) package

### As `SingleCellExperiment`  

```{r loom}  
loom <- example_loom()
## In memory
print(loom)
sce <- ingest_data(obj=loom) 

## From disk
print(loom$filename)
sce <- ingest_data(obj=loom$filename)
```

### As `Seurat`  

```{r} 
loom <- example_loom()
## In memory 
seurat <- ingest_data(obj=loom, 
                      output_type = "Seurat") 

## From disk
print(loom$filename)
seurat <- ingest_data(obj=loom$filename,
                      output_type = "Seurat")
```


# Cleanup 

Remove example files 

```{r, warning=F, message=F}
try({ file.remove("./pbmc_small.h5Seurat", showWarnings=F) })
try({ unlink("./pbmc_small_h5/", recursive = T) })
try({ file.remove("./pbmc_small.h5ad") })
try({ file.remove("./pbmc_small.loom")  })
```


# Session Info

<details>

```{r Session Info}
utils::sessionInfo()
```

</details>
