---
title: "Get started"
author: "<h5>Author: <i>Brian M. Schilder</i></h5>"  
date: "<h5>Most recent update: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h5>" 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
--- 

```{r setup}
library(scKirby)
```

# Intro 

`ingest_data()` runs three main steps:

1. **Read**: Automatically infers the file/object type and loads it (sub-function: `read_data()`).  
2. **Convert**: Converts it to the desired file/object type (sub-function: `to_<format>`). 
3. **Save**: Saves the converted file/object (sub-function: `save_data()`).

# Examples 

```{r}
seurat <- example_obj("seurat")
h5seurat <- example_obj("h5seurat")
sce <- example_obj("sce")
se <- example_obj("se")
h5 <- example_obj("h5")
cds <- example_obj("cds")
ctd <- example_obj("ctd")
adat <- example_obj("anndata")
library(Seurat)
loom <- example_obj("loom")
```


## Ingest expression matrix  

### As `SingleCellExperiment`  

```{r expression matrix}
sce2 <- ingest_data(obj = seurat, 
                    output_type = "SingleCellExperiment")
```

### As `Seurat` 

```{r}
seurat <- ingest_data(obj=seurat@assays$RNA@counts, 
                      output_type = "Seurat") 
```



## Ingest list 

`scKirby` can ingest a named list (i.e. `list(exp=..., annot=...)`) with the following items:  

- `exp`: Expression matrix with *rows/genes x cols/cells*. 
Can be a variety of matrix classes, including dense or sparse.  

- `annot`: Cell annotation `data.frame` with one cell per row. 
`rownames(annot)` should be the same as `colnames(exp)`. 

This happens to be the format that the example data in [`EWCE`](https://github.com/NathanSkene/EWCE) uses, but any user-supplied data will work. 

### As `SingleCellExperiment`  

```{r EWCElist}  
sce <- ingest_data(obj=ctd)
```

### As `Seurat` 

```{r} 
seurat.2 <- ingest_data(obj = ctd, 
                        output_type = "seurat")
```



## Ingest Seurat 

### As `SingleCellExperiment`  

In-memory

```{r Seurat}
sce.2 <- ingest_data(obj=seurat)
```

### Ingest H5Seurat  

On-disk 

```{r H5Seurat}   

seurat.2 <- ingest_data(obj = h5seurat, 
                        output_class = "seurat")
```


## Ingest HDF5 SingleCellExperiment 



### As `SingleCellExperiment`  

```{r HDF5}   
## Read in the sce object directly
h5.2 <- ingest_data(obj=h5)

## Read it from disk
sce_dir <- dirname(se_filepath(h5.2))
h5.3 <- ingest_data(obj=sce_dir)
```

### As `Seurat` 

```{r} 
## Read in the sce object directly
h5.4 <- ingest_data(obj = h5,
                    output_type = "Seurat") 
```

## Ingest CellDataSet 

```{r}  
seurat.2 <- ingest_data(obj=cds,
                        output_type = "Seurat") 
```



## Ingest AnnData 

### As `SingleCellExperiment`  

```{r AnnData}
# ### Set condaenv= to the name of a conda env you've made 
# reticulate::use_condaenv(condaenv = "echoR") 
## In memory
sce.2 <- ingest_data(obj=adat)

## On disk
path <- save_data(obj = adat,
                 filetype = "anndata")
sce.3 <- ingest_data(obj = path)
```

### As `Seurat` 

```{r}
seurat.2 <- ingest_data(obj=adat, 
                      output_type = "Seurat")
```


## Ingest loom

From [loomR](https://github.com/mojaveazure/loomR) package

### As `SingleCellExperiment`  

```{r loom}   
## In memory 
sce2 <- ingest_data(obj=loom) 

## From disk 
sce3 <- ingest_data(obj=loom$filename)
```

### As `Seurat`  

```{r}  
## In memory 
seurat2 <- ingest_data(obj=loom, 
                       output_type = "Seurat") 

## From disk
print(loom$filename)
seurat3 <- ingest_data(obj=loom$filename,
                       output_type = "Seurat")
```


# Cleanup 

Remove example files 

```{r, warning=F, message=F}
try({ file.remove("./pbmc_small.h5Seurat", showWarnings=FALSE) })
try({ unlink("./pbmc_small_h5/", recursive = TRUE) })
try({ file.remove("./pbmc_small.h5ad") })
try({ file.remove("./pbmc_small.loom")  })
```


# Session Info

<details>

```{r Session Info}
utils::sessionInfo()
```

</details>
<hr>
